Equations & R Code

### 2016-11-08 manuscript/coexistence-stability_ms.Rmd
updates storage effect theory section

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\allsectionsfont{\normalfont\sffamily\bfseries}

@@

\textsf{\large{\textbf{How fluctuation-dependent species coexistence affects the diversity-stability relationship}}}

@@

Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}

@@

#### Implementing the Storage Effect
For the storage effect to operate, we need species-specific responses to environmental variability, positive environment-competition (EC) covariance, and subadditive population growth [@Chesson1994; @Chesson2000a]. 
EC covariance limits population growth of residents: in a favorable year, they are limited by strong intraspecific competition because their populations are large. 
However, a species at very low abundance can escape the limitations of EC covariance if it experiences favorable conditions when the residents it is competiting against experience (relatively) unfavorable conditions. 
Subadditive population growth buffers populations against large population decreases in unfavorable years.

Our model includes subadditive population growth because we include a dormant stage with very low death rates, which limits large population declines.
Our model includes EC covariance through the effect of the dormant-to-live transition rate, $\gamma$, on population growth.
A high $\gamma$ year for a species represents a good *E* year in which live biomass, and thus resource uptake, is above average. 
These conditions correspond to *C* also being above average because a common species, by definition, cannot escape intraspecific competition.

In our model, rare species are rare because they are inferior competitors. 
So, given adequate variability in $\gamma$, the inferior competitor can persist because its rarity allows it to take advantage of favorable envrionmental conditions. 
We created competitive hierarchies in the storage effect version of the model by altering species' biomass conversion efficiencies ($\epsilon$)

@@

#### Implementing Relative Nonlinearity
\textcolor{red}{YOU NEED TO REMIND READERS ABOUT THE STANDARD RESOURCE COMPETITION MODELS, MACARTHUR, TILMAN, ETC. IN GENERAL, I THINK THIS PARAGRAPH NEEDS TO BE EXPANDED AND EXPLAINED MORE SLOWLY, STEP BY STEP.}

>>>>

\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}

<!--
\allsectionsfont{\normalfont\sffamily\bfseries}
-->

@@

\Large{How fluctuation-dependent species coexistence affects the diversity-stability relationship}

@@

\normalsize{Andrew T. Tredennick\textsuperscript{1}, Peter B. Adler\textsuperscript{1}, and Frederick R. Adler\textsuperscript{2}}

@@

#### Implementing the Storage Effect
For the storage effect to operate, we need species-specific responses to environmental variability, density-dependent covariance between environmental conditions and competition (*EC* covariance), and subadditive population growth [@Chesson1994; @Chesson2000a]. 
Regardless of the mechanism, long-term coexistence is possible when all species can increase when rare.
In the storage effect, rare species increase by escaping the effects of *EC* covariance.
This happens because common species will experience greater than average competition (*C*) in good environment (*E*) years because common species cannot avoid intraspecific competition.
Rare species do not have this problem and can increase rapdily in a good *E* year.
*EC* covariance is included in our model because dormant-to-live transition rates ($\gamma$) are species-specific and vary through time.
In a high $\gamma$ year for a common species, resource uptake will be above average, while in a high $\gamma$ year for a rare species, resource uptake will be below average.

Subadditive population growth buffers populations against large population decreases in unfavorable years. 
It is included in our model because we include a dormant stage with very low death rates, which limits large population declines.
In combination, subadditive population growth limits population declines in bad *E* years, and *EC* covariance ensures species can increase rapidly when rare but suffer when common.

@@

#### Implementing Relative Nonlinearity

\textcolor{red}{YOU NEED TO REMIND READERS ABOUT THE STANDARD RESOURCE COMPETITION MODELS, MACARTHUR, TILMAN, ETC. IN GENERAL, I THINK THIS PARAGRAPH NEEDS TO BE EXPANDED AND EXPLAINED MORE SLOWLY, STEP BY STEP.}

### 2016-11-08 manuscript/coexistence-stability_ms.tex
updates storage effect theory section

\allsectionsfont{\normalfont\sffamily\bfseries}

@@

For the storage effect to operate, we need species-specific responses to
environmental variability, positive environment-competition (EC)
covariance, and subadditive population growth (Chesson 1994, 2000b). EC
covariance limits population growth of residents: in a favorable year,
they are limited by strong intraspecific competition because their
populations are large. However, a species at very low abundance can
escape the limitations of EC covariance if it experiences favorable
conditions when the residents it is competiting against experience
(relatively) unfavorable conditions. Subadditive population growth
buffers populations against large population decreases in unfavorable
years.

Our model includes subadditive population growth because we include a
dormant stage with very low death rates, which limits large population
declines. Our model includes EC covariance through the effect of the
dormant-to-live transition rate, \(\gamma\), on population growth. A
high \(\gamma\) year for a species represents a good \emph{E} year in
which live biomass, and thus resource uptake, is above average. These
conditions correspond to \emph{C} also being above average because a
common species, by definition, cannot escape intraspecific competition.

In our model, rare species are rare because they are inferior
competitors. So, given adequate variability in \(\gamma\), the inferior
competitor can persist because its rarity allows it to take advantage of
favorable envrionmental conditions. We created competitive hierarchies
in the storage effect version of the model by altering species' biomass
conversion efficiencies (\(\epsilon\))

>>>>

@@

For the storage effect to operate, we need species-specific responses to
environmental variability, density-dependent covariance between
environmental conditions and competition (\emph{EC} covariance), and
subadditive population growth (Chesson 1994, 2000b). Regardless of the
mechanism, long-term coexistence is possible when all species can
increase when rare. In the storage effect, rare species increase by
escaping the effects of \emph{EC} covariance. This happens because
common species will experience greater than average competition
(\emph{C}) in good environment (\emph{E}) years because common species
cannot avoid intraspecific competition. Rare species do not have this
problem and can increase rapdily in a good \emph{E} year. \emph{EC}
covariance is included in our model because dormant-to-live transition
rates (\(\gamma\)) are species-specific and vary through time. In a high
\(\gamma\) year for a common species, resource uptake will be above
average, while in a high \(\gamma\) year for a rare species, resource
uptake will be below average.

Subadditive population growth buffers populations against large
population decreases in unfavorable years. It is included in our model
because we include a dormant stage with very low death rates, which
limits large population declines. In combination, subadditive population
growth limits population declines in bad \emph{E} years, and \emph{EC}
covariance ensures species can increase rapidly when rare but suffer
when common.
